Universal Anonymous Signatures

### 2022-10-04 prism-backend/research/gsac/formal-uas.tex
Spellchecked full paper.

\section{Formalizing \UAS}
\label{sec:formal-uas}

@@

Next we formalize our notion of Universal Anonymous Signatures
(\UAS). We define its syntax, model, and give a generic construction, which
we prove secure.

@@

\begin{description}
\item[$\Setup(\secpar) \rightarrow \parm$.] Given a security parameter \secpar,
  returns a global system parameter variable \parm. We assume that \parm are
  available to all other functions.
\item[$\KeyGen(\parm) \rightarrow (% \Ccom,r,
  \upk,\usk)$.] Given \parm, a user
  runs \KeyGen to generate its user key pair% , and a commitment \Ccom to the
  % public key, with randomness $r$
  . 
\item[$\OKeyGen(\parm,\finsp) \rightarrow (\opk,\osk)$.] Given \parm, and
  function \finsp, an opener runs \OKeyGen to generate its key pair. \finsp
  defines the utility extractable from signatures.
\item[$\ISet(\upk,\usk,\fissue) \rightarrow \ipk$.] Given a user key
  pair, and an issuance function \fissue for checking that credential requestors
  meet some conditions to be issued a credential, an existing user runs \ISet to
  extend the user public key, enabling this user to issue credentials meeting
  \fissue.
%\item[$\UKeyGen(\parm) \rightarrow (\upk,\usk)$.] Given \parm, returns a user's
%  key pair.
\item[$\langle
  \Obtain(\usk,\ipk,\sCred,\attrs),
  \Issue(\isk,\sipk,\attrs))  
  \rangle \rightarrow \langle \Cred/\bot,R/\bot
  \rangle$.] %
  This interactive protocol lets a user with key \usk receive a credential
  $\Cred=(\cid,\attrs,\cred,\ipk)$
  from an issuer in the system. $\cid=(\cidi,\cidu)$ is a unique identifier for
  the credential, where issuer and user randomly pick \cidi and \cidu,
  respectively (to ease notation, whenever we write just \cid, we mean $(\cidi,
  \cidu)$). \ipk is the issuer's public key and \cred is the actual credential,
  on attribute set \attrs. The user may leverage a set of previously obtained
  credentials $\sCred=\lbrace (\cid_i=(\cidi_i,\cidu_i),\cattrs_i,\cred_i,
  \ipk_i)\rbrace_{i\in[n]}$, from which we may omit the $\ipk_i$s for
  readability. The user outputs the produced credential \Cred, while the issuer
  outputs $R \gets (\yissue,\utrans)$, for some value $\yissue$ output by
  the \fissue function defined by the issuer, and $\utrans=(U,I)$ being the
  transcript of the communication% , where $U$ contains the messages sent by
  % the user, and $I$ those sent by the issuer
  .
\item[$\Sign(\usk,\opk,\sCred,\msg,\feval) \rightarrow (\sig,\yeval)$.] %
  Upon receiving user secret key \usk, opener public key \opk, 
  credentials \sCred, message \msg and evaluation function \feval, returns
  signature \sig, and a value \yeval. We use use \Sig to denote the
  tuple $(\sig,\yeval)$.
\item[$\Verify(\opk,\sipk,\Sig,\msg,\feval) \rightarrow 1/0$.]
  Checks whether $\Sig = (\sig,\yeval)$ is a valid signature
  over message \msg, from a user with credentials issued by issuers with public
  keys in \sipk, for evaluation function \feval and opener key \opk.
\item[$\Open(\osk,\sipk,\Sig,\msg,\feval) \rightarrow
  (\yinsp,\iproof)/\bot$.]
  Run by the opener with private key \osk. Receives a signature $\Sig=
  (\sig,\yeval)$ over message \msg and evaluation function \feval,
  generated using credentials by issuers with public keys in \sipk.
  If \Sig is valid, the function outputs a value $\yinsp$, and a proof of
  correct opening \iproof.
\item[$\Judge(\opk,\sipk,\yinsp,\iproof,\Sig,\msg,\feval) \rightarrow 1/0$.] %
  Checks if \iproof is a valid opening correctness proof for the value \yinsp,
  obtained by applying \Open to the the signature $\Sig = (\sig,\yeval)$
  over message \msg, and for evaluation function \feval. 
\end{description}

@@

Next, we define the concept of \CUASGen restrictions. These are special
cases of our generic construction that target a concrete variant of the
utility-privacy tradeoff by defining concrete \fissue, \feval and \finsp
functions. Security of \CUASGen implies security of its restrictions, if
the chosen \fissue function is \CUASGen-acceptable.

@@

\paragraph{Global Variables.} %
Users are referred to with user identifiers, \uid; for credentials, we use \cid;
and for openers, \oid. Although issuers are ``extended'' users, and thus can be
identified by their \uid, we mostly use \iid to refer to a user acting as
issuer, especially for clarity in operations involving both a user and an
issuer.
%
For notational convenience, we often use these identifiers instead of the
corresponding user/issuer/opener keys.
%
All tables/sets are initialized as empty tables/sets. We use the following sets
and tables in our definitions:

@@

\paragraph{Oracles.} %
We leverage the following oracles, formalized in \figref{fig:oracles1} and
\figref{fig:oracles2}.

\begin{description}
% \item[\IGEN.] Adds a new issuer to the game, generating its keypair and setting
%   the associated issuance function.
\item[\HUGEN.] Adds a new honest user, generating its keypair.
\item[\CUGEN.] Adds a new corrupt user. Sets its \upk to the
  one given by the adversary.  
\item[\OGEN.] Adds a new opener, generates its key pair and sets its opening
  function.
% \item[\ICORR.] Corrupts an existing (and honest) issuer, by giving its secret
%   key to the adversary.
\item[\ISET.] ``Upgrades'' a user to become an issuer, specifying its
  \fissue function.
\item[\UCORR.] Corrupts an existing honest user, giving its \usk to the
  adversary. If the user was also an issuer, it becomes a corrupt issuer too.
\item[\OCORR.] Corrupts an existing honest opener, giving its \osk to the
  adversary.
% \item[\HUGEN.] Adds a new honest user to the game, by honestly generating
%   the user's key pair.
% \item[\CUGEN.] Adds a new corrupt user to the game or, if the specified
%   user already exists and is honest, corrupts it, leaking its key and
%   credentials.
\item[\RREG.] Reads the given transcript table entry.
\item[\WREG.] Sets a transcript table entry to the given value.
\item[\OBTISS.] Lets the adversary add a new honestly generated credential to
  the game, on behalf of an honest user.
\item[\OBTAIN.] Enables the adversary to play the role of a dishonest issuer, by
  letting it interact with honest users who want to receive credentials.
\item[\ISSUE.] Allows the adversary to play the role of dishonest users,
  requesting an honest issuer to produce credentials for them.
\item[\SIGN.] Lets the adversary get signatures using credentials from honest
  users.
\item[\OPEN.] Given an honestly produced signature, outputs the result of the
  opening function, along with a correctness proof.
\item[\OBTCHALb.] Upon receiving a target credential identifier, two challenge
  users and credential sets, an honest issuer identifier, and a set of
  attributes, issues a credential for the challenge user and credential set
  defined by the bit $b$. 
\item[\CHALb.] Upon receiving two challenge users and credential sets, a common
  singing evaluation function and a message, returns a signature produced by one
  of these two user and credential sets, defined by the bit $b$.
\end{description}

@@

\paragraph{Helper functions % \SimSetup, \ExtractIssue, and \ExtractSign
  .}
We assume the existence of these helper functions. They are not functions
available in the actual scheme, but rather to the challenger in the experiments
we use to formalize security of \UAS schemes. We introduce them here, and define
them later for our construction.

@@

\paragraph{Correctness.} %
Correctness of \UAS schemes is formalized through the experiment in
\figref{fig:exp-uas-corr}. It states that a signature over any arbitrary message
and valid function \feval, produced honestly leveraging credential set \scid,
owned by user \uid, is accepted by \Verify. Moreover, all the credentials in
\scid meet the conditions set by the corresponding \fissue defined by the issuer
which issued each credential. Similarly, the output of \feval matches the
value produced by \Sign alongside with \sig; and the value produced by \Open
is accepted by \Judge, and matches the output of applying \finsp on \Yeval, the
credentials, user key, and message.

@@

Issuance anonymity captures that, during credential issuance
interactions, the adversary gains no information about the user(s) behind them,
beyond what is leaked by the policies required for issuance. For this, we allow
the adversary to control targetted issuers and openers, and add honest and
corrupt users. The adversary can also obtain credentials, produce signatures,
and open them. To prevent trivial wins: (1) when querying the oracle for
obtaining a challenge credential out of two possible user-credential set pairs,
the issuance protocol must either succeed or fail for both; and (2), the
adversary cannot mix challenge and non-challenge credentials in the same call to
the \SIGN oracle -- but can otherwise open any signature. With these
restrictions, if the adversary cannot distinguish arbitrary runs with all
oracles when he gets challenge credentials $\cred_b$, from runs with credentials
$\cred_{1-b}$, then the issuance protocol is privacy-preserving.

@@

We could simply check that all the credetnials involved in valid signatures
satisfy their respective \fissue functions, and belong to known users. However,
allowing credentials to endorse further credential requests makes this argument
insufficient. For instance, if we only check the credentials involved in
signature generation, it is perfectly possible that a forged credential is used
to obtain a legitimate one, and then use the latter to finally create a valid
signature -- concretly, such misbehaviour is not prevented by \ExpForgeSign.

@@

We formalize both properties in \figref{fig:exp-uas-unfor}.

@@

We formalize non-frameability in \ExpNonframe in \figref{fig:exp-uas-frame}.
The adversary is challenged to produce a signature
and opening proof that is accepted by \Verify and \Judge, respectively. From
the signature, we then extract the secret data used to generate it. The
adversary wins if the signature and opening proof are accepted, but the \yeval
or \yinsp values do not match what they should be based on the extracted data;
or if the signature can be associated to the secret key of an honest user who
did not produce it (i.e., no matching call to \SIGN exists).

@@

We give a generic construction of an \UAS scheme, based on
generic building blocks. We use three different NP relations: \RelIss, \RelSig,
and \RelIns, described next and specified in CS notation \cite{cs97} in
\figref{fig:nizkrels}. These relations include verifying that the comptuation of
the \fissue, \feval and \finsp functions has been done correctly.

>>>>

\section{Formalising \UAS}
\label{sec:formal-uas}

@@

Next we formalise our notion of Universal Anonymous Signatures
(\UAS). We define its syntax, model, and give a generic construction, which
we prove secure.

@@

\begin{description}
\item[$\Setup(\secpar) \rightarrow \parm$.] Given a security parameter \secpar,
  returns a global system parameter variable \parm. We assume that \parm are
  available to all other functions.
\item[$\KeyGen(\parm) \rightarrow (% \Ccom,r,
  \upk,\usk)$.] Given \parm, a user
  runs \KeyGen to generate its user key pair% , and a commitment \Ccom to the
  % public key, with randomness $r$
  . 
\item[$\OKeyGen(\parm,\finsp) \rightarrow (\opk,\osk)$.] Given \parm, and
  function \finsp, an opener runs \OKeyGen to generate its key pair. \finsp
  defines the utility extractable from signatures.
\item[$\ISet(\upk,\usk,\fissue) \rightarrow \ipk$.] Given a user key
  pair, and an issuance function \fissue for checking that credential requesters
  meet some conditions to be issued a credential, an existing user runs \ISet to
  extend the user public key, enabling this user to issue credentials meeting
  \fissue.
%\item[$\UKeyGen(\parm) \rightarrow (\upk,\usk)$.] Given \parm, returns a user's
%  key pair.
\item[$\langle
  \Obtain(\usk,\ipk,\sCred,\attrs),
  \Issue(\isk,\sipk,\attrs))  
  \rangle \rightarrow \langle \Cred/\bot,R/\bot
  \rangle$.] %
  This interactive protocol lets a user with key \usk receive a credential
  $\Cred=(\cid,\attrs,\cred,\ipk)$
  from an issuer in the system. $\cid=(\cidi,\cidu)$ is a unique identifier for
  the credential, where issuer and user randomly pick \cidi and \cidu,
  respectively (to ease notation, whenever we write just \cid, we mean $(\cidi,
  \cidu)$). \ipk is the issuer's public key and \cred is the actual credential,
  on attribute set \attrs. The user may leverage a set of previously obtained
  credentials $\sCred=\lbrace (\cid_i=(\cidi_i,\cidu_i),\cattrs_i,\cred_i,
  \ipk_i)\rbrace_{i\in[n]}$, from which we may omit the $\ipk_i$s for
  readability. The user outputs the produced credential \Cred, while the issuer
  outputs $R \gets (\yissue,\utrans)$, for some value $\yissue$ output by
  the \fissue function defined by the issuer, and $\utrans=(U,I)$ being the
  transcript of the communication% , where $U$ contains the messages sent by
  % the user, and $I$ those sent by the issuer
  .
\item[$\Sign(\usk,\opk,\sCred,\msg,\feval) \rightarrow (\sig,\yeval)$.] %
  Upon receiving user secret key \usk, opener public key \opk, 
  credentials \sCred, message \msg and evaluation function \feval, returns
  signature \sig, and a value \yeval. We use use \Sig to denote the
  tuple $(\sig,\yeval)$.
\item[$\Verify(\opk,\sipk,\Sig,\msg,\feval) \rightarrow 1/0$.]
  Checks whether $\Sig = (\sig,\yeval)$ is a valid signature
  over message \msg, from a user with credentials issued by issuers with public
  keys in \sipk, for evaluation function \feval and opener key \opk.
\item[$\Open(\osk,\sipk,\Sig,\msg,\feval) \rightarrow
  (\yinsp,\iproof)/\bot$.]
  Run by the opener with private key \osk. Receives a signature $\Sig=
  (\sig,\yeval)$ over message \msg and evaluation function \feval,
  generated using credentials by issuers with public keys in \sipk.
  If \Sig is valid, the function outputs a value $\yinsp$, and a proof of
  correct opening \iproof.
\item[$\Judge(\opk,\sipk,\yinsp,\iproof,\Sig,\msg,\feval) \rightarrow 1/0$.] %
  Checks if \iproof is a valid opening correctness proof for the value \yinsp,
  obtained by applying \Open to the the signature $\Sig = (\sig,\yeval)$
  over message \msg, and for evaluation function \feval. 
\end{description}

@@

Next, we define the concept of \CUASGen restrictions. These are special
cases of our generic construction that target a concrete variant of the
utility-privacy trade-off by defining concrete \fissue, \feval and \finsp
functions. Security of \CUASGen implies security of its restrictions, if
the chosen \fissue function is \CUASGen-acceptable.

@@

\paragraph{Global Variables.} %
Users are referred to with user identifiers, \uid; for credentials, we use \cid;
and for openers, \oid. Although issuers are ``extended'' users, and thus can be
identified by their \uid, we mostly use \iid to refer to a user acting as
issuer, especially for clarity in operations involving both a user and an
issuer.
%
For notational convenience, we often use these identifiers instead of the
corresponding user/issuer/opener keys.
%
All tables/sets are initialised as empty tables/sets. We use the following sets
and tables in our definitions:

@@

\paragraph{Oracles.} %
We leverage the following oracles, formalised in \figref{fig:oracles1} and
\figref{fig:oracles2}.

\begin{description}
% \item[\IGEN.] Adds a new issuer to the game, generating its keypair and setting
%   the associated issuance function.
\item[\HUGEN.] Adds a new honest user, generating its key pair.
\item[\CUGEN.] Adds a new corrupt user. Sets its \upk to the
  one given by the adversary.  
\item[\OGEN.] Adds a new opener, generates its key pair and sets its opening
  function.
% \item[\ICORR.] Corrupts an existing (and honest) issuer, by giving its secret
%   key to the adversary.
\item[\ISET.] ``Upgrades'' a user to become an issuer, specifying its
  \fissue function.
\item[\UCORR.] Corrupts an existing honest user, giving its \usk to the
  adversary. If the user was also an issuer, it becomes a corrupt issuer too.
\item[\OCORR.] Corrupts an existing honest opener, giving its \osk to the
  adversary.
% \item[\HUGEN.] Adds a new honest user to the game, by honestly generating
%   the user's key pair.
% \item[\CUGEN.] Adds a new corrupt user to the game or, if the specified
%   user already exists and is honest, corrupts it, leaking its key and
%   credentials.
\item[\RREG.] Reads the given transcript table entry.
\item[\WREG.] Sets a transcript table entry to the given value.
\item[\OBTISS.] Lets the adversary add a new honestly generated credential to
  the game, on behalf of an honest user.
\item[\OBTAIN.] Enables the adversary to play the role of a dishonest issuer, by
  letting it interact with honest users who want to receive credentials.
\item[\ISSUE.] Allows the adversary to play the role of dishonest users,
  requesting an honest issuer to produce credentials for them.
\item[\SIGN.] Lets the adversary get signatures using credentials from honest
  users.
\item[\OPEN.] Given an honestly produced signature, outputs the result of the
  opening function, along with a correctness proof.
\item[\OBTCHALb.] Upon receiving a target credential identifier, two challenge
  users and credential sets, an honest issuer identifier, and a set of
  attributes, issues a credential for the challenge user and credential set
  defined by the bit $b$. 
\item[\CHALb.] Upon receiving two challenge users and credential sets, a common
  singing evaluation function and a message, returns a signature produced by one
  of these two user and credential sets, defined by the bit $b$.
\end{description}

@@

\paragraph{Helper functions % \SimSetup, \ExtractIssue, and \ExtractSign
  .}
We assume the existence of these helper functions. They are not functions
available in the actual scheme, but rather to the challenger in the experiments
we use to formalise security of \UAS schemes. We introduce them here, and define
them later for our construction.

@@

\paragraph{Correctness.} %
Correctness of \UAS schemes is formalised through the experiment in
\figref{fig:exp-uas-corr}. It states that a signature over any arbitrary message
and valid function \feval, produced honestly leveraging credential set \scid,
owned by user \uid, is accepted by \Verify. Moreover, all the credentials in
\scid meet the conditions set by the corresponding \fissue defined by the issuer
which issued each credential. Similarly, the output of \feval matches the
value produced by \Sign alongside with \sig; and the value produced by \Open
is accepted by \Judge, and matches the output of applying \finsp on \Yeval, the
credentials, user key, and message.

@@

Issuance anonymity captures that, during credential issuance
interactions, the adversary gains no information about the user(s) behind them,
beyond what is leaked by the policies required for issuance. For this, we allow
the adversary to control targeted issuers and openers, and add honest and
corrupt users. The adversary can also obtain credentials, produce signatures,
and open them. To prevent trivial wins: (1) when querying the oracle for
obtaining a challenge credential out of two possible user-credential set pairs,
the issuance protocol must either succeed or fail for both; and (2), the
adversary cannot mix challenge and non-challenge credentials in the same call to
the \SIGN oracle -- but can otherwise open any signature. With these
restrictions, if the adversary cannot distinguish arbitrary runs with all
oracles when he gets challenge credentials $\cred_b$, from runs with credentials
$\cred_{1-b}$, then the issuance protocol is privacy-preserving.

@@

We could simply check that all the credentials involved in valid signatures
satisfy their respective \fissue functions, and belong to known users. However,
allowing credentials to endorse further credential requests makes this argument
insufficient. For instance, if we only check the credentials involved in
signature generation, it is perfectly possible that a forged credential is used
to obtain a legitimate one, and then use the latter to finally create a valid
signature -- concretely, such misbehaviour is not prevented by \ExpForgeSign.

@@

We formalise both properties in \figref{fig:exp-uas-unfor}.

@@

We formalise non-frameability in \ExpNonframe in \figref{fig:exp-uas-frame}.
The adversary is challenged to produce a signature
and opening proof that is accepted by \Verify and \Judge, respectively. From
the signature, we then extract the secret data used to generate it. The
adversary wins if the signature and opening proof are accepted, but the \yeval
or \yinsp values do not match what they should be based on the extracted data;
or if the signature can be associated to the secret key of an honest user who
did not produce it (i.e., no matching call to \SIGN exists).

@@

We give a generic construction of an \UAS scheme, based on
generic building blocks. We use three different NP relations: \RelIss, \RelSig,
and \RelIns, described next and specified in CS notation \cite{cs97} in
\figref{fig:nizkrels}. These relations include verifying that the computation of
the \fissue, \feval and \finsp functions has been done correctly.
